---
title: "integrated_3_datasets_epithelial"
author: "SandraTietscher"
date: "2022-05-15"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Load packages

```{r load-packages}
library(Seurat)
library(ggplot2)
library(RColorBrewer)
library(ggpubr)
```

# Read in data

```{r read-data}
tietscher <- readRDS("data/final_seurat_object_filtered_noSCT.rds")
bassez <- readRDS("data/Bassez2021_scRNAseq_data/Bassez_cohort1_seurat.rds")
qian <- readRDS("data/Qian2020_scRNAseq_data/qian_seurat.rds")
```

# Prepare datasets

```{r prepare-datasets}
# Add metadata column to indicate which dataset the cell belongs to
bassez$dataset <- "bassez"
tietscher$dataset <- "tietscher"
qian$dataset <- "qian"

# Exclude on-treatment samples from bassez dataset
bassez <- subset(bassez, timepoint == "Pre")

```

# Subset epithelial cells 

Randomly subsample each dataset to 10000 cells.

```{r}
# Subset cancer cells for both datasets
bassez_epi <- subset(bassez, cellType == "Cancer_cell")
tietscher_epi <- subset(tietscher, cell.type == "epithelial")
qian_epi <- subset(qian, CellType == "Cancer")

# Harmonize patient metadata
unique(bassez_epi$patient_id)
tietscher_epi$patient_id <- tietscher_epi$sample
qian_epi$patient_id <- paste0("QianPatient_", qian_epi$PatientNumber)

bassez_sub <- bassez_epi[, sample(colnames(bassez_epi), size = 10000, replace=F)]
tietscher_sub <- tietscher_epi[, sample(colnames(tietscher_epi), size = 10000, replace=F)]
qian_sub <- qian_epi[, sample(colnames(qian_epi), size = 10000, replace=F)]

datasets_epi <- list(tietscher_sub, bassez_sub, qian_sub)
```


# Normalize and identify variable features

```{r normalization}
# normalize and identify variable features for each dataset independently
datasets_epi <- lapply(X = datasets_epi, FUN = function(x) {
    x <- NormalizeData(x)
    x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 2500)
})

# select features that are repeatedly variable across datasets for integration
features <- SelectIntegrationFeatures(object.list = datasets_epi, nfeatures = 2500)
```

# Perform integration

We then identify anchors using the FindIntegrationAnchors() function, which takes a list of Seurat objects as input, and use these anchors to integrate the two datasets together with IntegrateData().

```{r integration}
integr.anchors <- FindIntegrationAnchors(object.list = datasets_epi, anchor.features = features)

# this command creates an 'integrated' data assay
datasets.combined <- IntegrateData(anchorset = integr.anchors)

# specify that we will perform downstream analysis on the corrected data. Note that the original unmodified data still resides in the 'RNA' assay
DefaultAssay(datasets.combined) <- "integrated"

saveRDS(datasets.combined, "data/epithelial_tietscher_bassez_qian_10000cells.rds")
```

# Average epithelial HLA-ABC expression per patient

```{r}
# Sum up integrated scaled counts for HLA-A, HLA-B and HLA-C
datasets.combined$HLAABC_counts <- colSums(datasets.combined@assays$integrated@data[c("HLA-A", "HLA-B"),])

# Average over patients and add dataset information
colData <- datasets.combined@meta.data
HLAABC_mean <- colData %>% group_by(patient_id) %>% dplyr::summarise(mean = mean(HLAABC_counts))
HLAABC_mean$dataset <- ifelse(grepl("BIOKEY", HLAABC_mean$patient_id), "bassez", ifelse(grepl("Qian", HLAABC_mean$patient_id), "qian", "tietscher"))

write.csv(HLAABC_mean, "output/epiHLAABC_patient_mean_3datasets.csv")
```

